### 2025-05-08 futureWork/rlHedging.tex
new

In the evolving landscape of financial markets, insurance products such as VAs have gained a substantial amount of interest due to their ability to provide both investment growth and guaranteed benefits. 
Managing the risks associated with these products, especially in volatile market conditions, is a complex task that demands sophisticated financial modeling techniques. 
\gls{rl} has emerged as a powerful tool for addressing such complexities. With the recent advancement in deep learning, \gls{dnn} architectures have been successfully applied to \gls{rl} problems, leading to the development of deep reinforcement learning (\gls{drl}) algorithms. 
These algorithms have demonstrated remarkable success decision-making tasks that must be optimized over time. 
In the context of managing VAs, \gls{drl} has shown promise in optimizing hedging strategies where long-term financial and policyholder decisions interact dynamically with changing market conditions.

@@

The rest of the chapter is organized as follows: 
Section~\ref{sec:MDP} introduces the mathematical formulation for hedging VAs and defines the key components of the framework.
Section~\ref{sec:RLAlgorithms} provides an overview of existing \gls{rl} algorithms and their limitations in the context of hedging VAs.
Section~\ref{sec:PPO} introduces the proximal policy optimization algorithm and its application to the hedging of VAs.
Finally, Section~\ref{sec:FutureDirections} concludes the chapter and discusses future research directions.

\section{Markov Decision Process for Hedging VAs} \label{sec:MDP}

@@

The objective in an \gls{mdp} is to find a policy, $\pi: \mathcal{S} \rightarrow \mathbb{P}(\mathcal{A})$, that maximizes the cumulative reward over time.
A \gls{mdp} provides a mathematical framework for solving sequential decision-making tasks in an enviornment where outcomes are partly random and partly under the control of a hedging agent. 
For hedging VAs, the asset dynamics and the mortality model are not controlled by the agent, but the agent controls the cumulative reward by deciding on the hedging weights.
In this section, we reformulate the hedging enviornment of VAs from Section~\ref{subsec:VApayout} as an \gls{mdp} and define the key components of the \gls{mdp}.

@@

At the inception of the contract, i.e., $t=0$, we assume that the whole premium is invested in the stock index and the guarantee base is set to the sub-account value:
\begin{equation*}
    S_0=F_0=G_0.
\end{equation*}
At each time $t=1,\ldots,T$, the following events take place in the following order:
\begin{enumerate}
    \item The dynamic lapse rate $q_t$ is applied to the contract, i.e., $q$ of the policyholders leave the contract at time $t$.
        \begin{equation} \label{eq3:lapse}
            q_t = q_t^B \cdot \text{clip} (M_q (\frac{G_{t-1}}{F_{t-1}} - D_q), L_q, U_q),
        \end{equation}
    where $q_t^B$ is the base lapse rate, $\text{clip}(x, a, b)$ is a function that clips the value of $x$ to be within the range $[a, b]$, and $M_q$, $D_q$, $L_q$, and $U_q$ are the model parameters taken from National Association of Insurance Commissioners’ (NAIC) Valuation Manual 21~\citep{naic2021}.
    \item The sub-account value changes according to the growth of the underlying stock and the (gross) management fee is deducted. That is, 
        \begin{equation} \label{eq3:subaccount}
            F_t = F_{(t-1)_+}\cdot\frac{S_{t}}{S_{t-1}}\cdot(1-\eta_g)\cdot(1-q_t),
        \end{equation} 
    where $(x)^+=\max\{x,0\}$ and $F_{(t-1)_+}$ will be defined later. The insurer's income at time $t$ is the net management fee, i.e., $F_t\eta_n$.

@@

Consider a \gls{va} contract whose delta hedge portfolio at any time~$t$, $t=0,1,\ldots,T-1$, consists of $\Delta_t$ units in the underlying stock and $B_t$ amount of a risk-free zero-coupon bond maturing at time $T$.
The value of the hedge portfolio at time~$(t-1)$ is:
\begin{equation*}
    H_{t-1} = \Delta_{t-1} S_{t-1} + B_{t-1},
\end{equation*}
where $S_t$ is the underlying stock price and any time $t>0$.
This hedge portfolio is brought forward to the next rebalancing time~$t$, when its value becomes:
\begin{equation*}
    H_{t}^{bf} = \Delta_{t-1} S_{t} + B_{t-1}e^{-rt}.
\end{equation*}
Therefore, the time~$t$ hedging error, i.e., the cash flow incurred by the insurer due to rebalancing at time~$t$, is
\begin{equation}
    HE_t = H_t - H^{bf}_t, \quad t=1,\ldots, T-1.
\end{equation}
The P\&L of the \gls{va} contract includes the cost of the initial hedge ($H_0$), the hedging errors~\eqref{eq2:hedgingerror}, the unwinding of the hedge at maturity ($H^{bf}_T$), and the contract payout at time~$t\in \{0,\ldots,T\}$.
Mathematically, the present value of the liability for a \gls{gmmb} contract is 
\begin{align} \label{eq3:lossGMMB}
L   & = H_0 - e^{-rT} H^{bf}_T + \sum_{t=1}^{T-1} e^{-rt} HE_t + \sum_{t=1}^T e^{-rt} C_t - \sum_{t=1}^T e^{-rt} F_t\eta_n + e^{-rT} (G_t - F_t)^+  \nonumber \\ 
    & = e^{-rT} (G_T - F_{T^+})^+ + \sum_{t=1}^T e^{-rt}  \left( \Delta_{t-1} (S_{t-1} - e^{-r} S_t) + C_t - F_t\eta_n \right) 
\end{align}
where the equality holds by a rearrangement of terms and a telescopic sum simplification of $e^{-rt}B_t$, $t=0,\ldots,T-1$, and $C_t := C \cdot S_{t-1} \cdot |\Delta_t - \Delta_{t-1}|$ is the tranaction cost at time $t$~\citep{garleanu2013dynamic}.
Hence, the present liability consist of the hedging errors, the transaction costs, the net management fees, and the contract payouts.
Similarly, the present value of the liability for a \gls{gmwb} contract is
\begin{equation} \label{eq3:lossGMWB}
L = \sum_{t=1}^T e^{-rt}  \left( \Delta_{t-1} (S_{t-1} - e^{-r} S_t) + C_t - F_t\eta_n \right)
\end{equation}
The \gls{gmwb} payout is implicitly included in $F_t\eta_n$.
Both Equation~\eqref{eq3:lossGMMB} and Equation~\eqref{eq3:lossGMWB} neglect the transaction cost to liquidate the hedge portfolio at maturity.
The discrete-time hedging problem of VAs can be conveniently formulated as an \gls{mdp}, where the states, actions, transition probabilities, policies, and rewards are defined as follows.
Different from the formulation in Chapter~\ref{chap:project2}, the hedging weight $\Delta_0, \ldots, \Delta_{T-1}$ are not estimated from inner simulation steps but are directly controlled by a \gls{drl} agent.
Thus, the hedging weight $\Delta_t$ at time $t$ is estimated using a neural network whose input is the state $\mathbf{s}_t$ and output is $\Delta_t$.

@@

For hedging VAs,~\cite{xu2020variable} propose a deep hedging algorithm that falls into this category of model-based \gls{rl}.
A cost function is defined to minimize the expected shortfall of the total present liability.
~\cite{carbonneau2021deep} extends the model-based approach to hedging long-term financial derivatives and compares the performance of deep hedging strategies trained with different cost functions.
In deep hedging~\citep{buehler2019deep} the state space of the \gls{mdp} contains the previous hedging weight.
A more recent paper by~\cite{imaki2021no} proposes a new network architecture that removes the previous hedging weight from the state space by introducing a no-transaction band.
It allows for a more efficient training of the neural network by removing the semi-recurrent problem structure.

@@

In a model-free reinforcement learning approach, the agent is a trial-and-error learner that interacts with the environment to learn the optimal policy by iteratively updating the policy and the value function based on the observed states, actions, and rewards.
For hedging VAs, the agent does not have access to a planning algorithm or the underlying stochastic asset model, and it must learn the optimal policy by directly interacting with the environment.

Model-free \gls{rl} algorithms gain its popularity in robotics tasks, where transition probabilities in the environment are unknown and can be affected by the agent's actions.
In the context of hedging, existing studies use transaction costs as a function of the hedging weights, which makes the transition probabilities independent on the actions.
However, the trade size can affect the asset price in a real environment~\citep{hasbrouck1991measuring}, which makes the model-free approach a promising direction for hedging VAs in the real market.

@@

Representative algorithms include the \gls{trpo} algorithm proposed by~\cite{schulman2015trust} and the \gls{ppo} algorithm proposed by~\cite{schulman2017proximal}.
~\cite{du2020deep} show that the \gls{ppo} algorithm is more sample efficient and stable than the value-based approaches in training agents to hedge financial derivatives.
The most relevant paper to our study is~\cite{chong2023pseudo}, which uses the \gls{ppo} algorithm to hedge \gls{gmmb} with a GMDB rider under the Black-Scholes framework.
However, information of the underlying risk factors is leaked to the agent, which makes their study pseudo-model-free.
To further improve the risk management of VAs under rough volatility models, we propose a model-free deep reinforcement learning approach to hedge VAs without explicit or implicit knowledge of the underlying risk factors.

\section{Hedging VAs with a PPO Agent} \label{sec:PPO}
In a hedging problem, an insurer aims at finding a policy $\pi$ that maximizes the expected reward over time.
In a policy-based \gls{rl} approach like the \gls{ppo}, both the policy and the value function are parameterized by deep neural networks.
We denote the policy by $\pi_{\theta}$ and the value function by $V_{\phi}$, where $\theta$ and $\phi$ are the parameters of the value network and the policy network, respectively.
The policy is updated according to an objective function that reflects the advantage of taking action $a_t$ in state $s_t$ over following the policy $\pi$.
\begin{align}
    \max_{\theta} \mathbb{E}\left[ \min \left\{ \Lambda_t(\theta)\hat{A}^{\pi_{\theta}}(s_t, a_t), \text{clip}(\Lambda_t(\theta), 1-\epsilon, 1 + \epsilon) \hat{A}^{\pi_{\theta}}(s_t, a_t)  \right\} \right] \\
    \Lambda_t(\theta) = \frac{\pi_{\theta}(a_t|s_t)}{\pi_{\theta_{old}}(a_t|s_t)} 
\end{align}
where $\theta_{old}$ denotes the policy parameters at the previous iteration, $\epsilon$ is a clip range hyperparameter that controls the size of the policy update, $\text{clip}(x, a, b)$ is a function that clips the value of $x$ to be within the range $[a, b]$, and $\hat{A}^{\pi_{\theta}}(s_t, a_t)$ is an estimated advantage of taking action $a_t$ in state $s_t$ over following the policy $\pi_{\theta_{old}}$.
In our implementation, the advantage function is estimated following the GAE method proposed by~\cite{schulman2015high}.
The policy is updated by performing gradient ascent with respect to the policy parameters $\theta$ to maximize the objective function, which is estimated by sampling from the environment and computing the empirical average of the objective function over the sampled trajectories.
The first component in the objective function, $\Lambda_t(\theta) \hat{A}^{\pi_{\theta}}(s_t, a_t)$, is a surrogate objective function that approximates the advantage of taking action $a_t$ in state $s_t$ over following the policy $\pi_{\theta_{old}}$, and the second component, $\text{clip}(\Lambda_t(\theta), 1-\epsilon, 1 + \epsilon) \hat{A}^{\pi_{\theta}}(s_t, a_t)$, is a clipped surrogate objective function that prevents large policy updates and ensures more stable training.
The \gls{trpo} algorithm proposed by~\cite{schulman2015trust} uses a similar clipped surrogate objective function to ensure that the policy update does not deviate too far from the old policy, but \gls{ppo} simplifies the optimization problem by using a clipped surrogate objective function instead of a hard-coded constraint on the policy update.
Compared to \gls{trpo}, \gls{ppo} is more computationally efficient and easier to implement, making it a popular choice for training deep reinforcement learning agents.
For a in-depth discussion on the implementation details of both \gls{trpo} and \gls{ppo}, we refer the reader to~\cite{engstrom2020implementation}.

@@

Algorithm~\ref{alg3:ppoHedging} outlines the \gls{ppo} algorithm for training an agent to hedge VAs.
At each iteration $k$, the agent collects a set of trajectories $\mathcal{D}_k$ by interacting with the environment (Section~\ref{subsec:VASimulation}) with the current policy $\pi_{\theta_k}$.
The agent then computes the rewards $\hat{R}_t$ for each trajectory in $\mathcal{D}_k$ and estimates the advantage function with $\hat{A}_{\pi_{\theta_k}}(s_t, a_t)$ using the value function $V_{\phi_k}$.
The policy parameters $\theta_{k+1}$ are updated by performing gradient ascent with respect to the policy parameters to maximize the clipped surrogate objective function.
The value function parameters $\phi_{k+1}$ are updated by minimizing the mean squared error between the predicted value and the actual collected reward.
The agent iterates this process until convergence to learn an optimal policy for hedging VAs.

@@

In the context of hedging VAs, the state space is partially observed.
The agent does not have access to the contract specifications and the underlying risk factors.
Furthermore, a rough volatility model introduces non-Markovian dynamics, i.e., the future state of the system depends on the entire history of the system.
To address the non-Markovian dynamics, an \gls{lstm} component is added to the state representation $\mathbf{s}_t$ to capture the temporal dependencies in the data.
This approach is known as recurrent \gls{ppo}~\citep{ni2021recurrent}.
Let $h_t$ be the hidden state of the \gls{lstm} at time $t$.
The state representation $\mathbf{s}_t$ is fed into the \gls{lstm} at each time step to update the hidden state $h_t$.
\begin{equation} \label{eq3:stateRNN}
    \mathbf{s}_t = (S_t, F_t, G_t, \tau, \Delta_{t-1}), \quad h_t = \text{\gls{lstm}}(\mathbf{s}_t, h_{t-1})
\end{equation}

In a \gls{lstm}, the cell state $c_t$ stores a representation of historical information up to time $t$.
It is updated at each time step by a forget gate $f_t$, an input gate $i_t$, and an output gate $o_t$.

In a recurrent \gls{ppo} with \gls{lstm}, the policy and reward function remain the same but with the new state representation in Equation~\eqref{eq3:stateRNN}.
Instead of following the standard \gls{ppo} algorithm, the agent uses the recurrent \gls{ppo} algorithm to learn an optimal policy for hedging VAs.
Figure~\ref{fig3:architecturePPO} illustrates the network architecture of the standard \gls{ppo} and the recurrent \gls{ppo} with \gls{lstm}, and Algorithm~\ref{alg3:ppoHedging-rnn} outlines the recurrent \gls{ppo} algorithm and for training an agent to hedge VAs with non-Markovian dynamics.

@@

\section{Transfer Learning for Hedging VAs} \label{sec:TransferLearning}

In the context of deep hedging, various research has been conducted to improve the hedging strategies on real-world data by leveraging knowledge from simulated data.
Most of the research uses the offline-online learning approach, where the hedging strategies are first trained with simulation data, but updated in real-time based on new market information and continuous learning to adapt to changing market dynamics.
The most relevant paper to our study is~\cite{xiao2021optimal}, which use policy gradient algorithms to train hedging strategies for European options with simulation data from Heston model and then updated in real-time hedging S\&P 500 index options.
To the best of our knowledge, no research has been conducted to improve the hedging strategies with the state-of-the-art transfer learning techniques.
This section introduces some transfer learning techniques that are relevant to our task of hedging VAs.

@@

In the context of hedging VAs using deep reinforcement learning and transfer learning, reward shaping becomes a crucial technique. 
It involves modifying the reward function in the target domain to incorporate additional guidance, which often leads to more efficient learning and a better performance.
This technique can be particularly useful for managing the complex risks associated with VAs, as it can guide the learning process towards strategies that are more effective in hedging these products.

@@

We present a numerical experiment to demonstrate the effectiveness of the model-free \gls{ppo} algorithm for hedging VAs with transaction costs.
We consider a \gls{va} contract with a maturity of $T=240$ months and a \gls{gmmb} rider with a simulation environment described in Section~\ref{subsec:VASimulation}.
The simulation enviornment is implemented in Python using the OpenAI Gym framework~\citep{brockman2016openai}, and the neural networks are programmed using the PyTorch library~\citep{paszke2019pytorch}.
The environment is designed to simulate the dynamics of the \gls{va} contract and provide the agent with the current state of the \gls{va} with the contract specifications and the financial market information described in Table~\ref{tab3:hyperparameters}.

@@

Figure~\ref{fig3:dh-transfers} illustrates the hedging performance of the deep hedging algorithm with transfer learning for hedging VAs with a \gls{gmmb} rider.
The first deep hedging agent is directly trained on the Heston asset model, and the second deep hedging agent is trained with transfer learning from the \gls{gbm} model.
The figure plot the 95\%-CVaR of hedging errors against training epochs for both agents.
The deep hedging agent with transfer learning outperforms the deep hedging agent trained on the Heston model in both mastery and generalization.
The agent with transfer learning achieves a lower loss in the target domain and adapts more quickly to the new environment, indicating that the agent has learned a more robust hedging strategy that can be effectively applied to a more complex environment.

@@

This experiment compares the hedging performance of the recurrent \gls{ppo} agent with \gls{lstm} and the deep hedging algorithm proposed by~\cite{buehler2019deep} for hedging VAs with a \gls{gmmb} rider.
The experiment is conducted with the same amount ofsimulation budget.

>>>>

In the evolving landscape of financial markets, insurance products such as \gls{va} contracts have gained a substantial amount of interest due to their ability to provide both investment growth and guaranteed benefits. 
Managing the risks associated with these products, especially in volatile market conditions, is a complex task that demands sophisticated financial modeling techniques. 
\gls{rl} has emerged as a powerful tool for addressing such complexities. With the recent advancement in deep learning, \gls{dnn} architectures have been successfully applied to \gls{rl} problems, leading to the development of \gls{drl} algorithms. 
These algorithms have demonstrated remarkable success decision-making tasks that must be optimized over time. 
In the context of managing \gls{va} contracts, \gls{drl} has shown promise in optimizing hedging strategies where long-term financial and policyholder decisions interact dynamically with changing market conditions.

@@

The rest of the chapter is organized as follows: 
Section~\ref{sec:MDP} introduces the mathematical formulation for hedging \gls{va} contracts and defines the key components of the framework.
Section~\ref{sec:RLAlgorithms} provides an overview of existing \gls{rl} algorithms and their limitations in the context of hedging \gls{va} contracts.
Section~\ref{sec:PPO} introduces the proximal policy optimization algorithm and its application to the hedging of \gls{va} contracts.
Finally, Section~\ref{sec:FutureDirections} concludes the chapter and discusses future research directions.

\section{Markov Decision Process for Hedging VA Contracts} \label{sec:MDP}

@@

The objective in an \gls{mdp} is to find a policy, $\pi: \mathcal{S} \rightarrow \mathbb{P}(\mathcal{A})$, that maximizes the cumulative reward over time.
A \gls{mdp} provides a mathematical framework for solving sequential decision-making tasks in an enviornment where outcomes are partly random and partly under the control of a hedging agent. 
For hedging \gls{va} contracts, the asset dynamics and the mortality model are not controlled by the agent, but the agent controls the cumulative reward by deciding on the hedging weights.
In this section, we reformulate the hedging enviornment of \gls{va} contracts from Section~\ref{subsec:VApayout} as an \gls{mdp} and define the key components of the \gls{mdp}.

@@

At the inception of the contract, i.e., $t=0$, we assume that the whole premium is invested in the stock index and the guarantee base is set to the sub-account value:
\begin{equation*}
    S_0=F_0=G_0.
\end{equation*}
At each time $t=1,\ldots,T$, the following events take place in the following order:
\begin{enumerate}
    \item The dynamic lapse rate $q_t$ is applied to the contract, i.e., $q$ of the policyholders leave the contract at time $t$.
        \begin{equation} \label{eq3:lapse}
            q_t = q_t^B \cdot \text{clip} (M_q (\frac{G_{t-1}}{F_{t-1}} - D_q), L_q, U_q),
        \end{equation}
    where $q_t^B$ is the base lapse rate, $\text{clip}(x, a, b)$ is a function that clips the value of $x$ to be within the range $[a, b]$, i.e.,
    $$
    \text{clip}(x, a, b) = \begin{cases}
        a, & \text{if } x < a \\
        x, & \text{if } a \leq x \leq b \\
        b, & \text{if } x > b,
    \end{cases}
    $$
    and $M_q$, $D_q$, $L_q$, and $U_q$ are the model parameters taken from National Association of Insurance Commissioners’ (NAIC) Valuation Manual 21~\citep{naic2021}.
    \item The sub-account value changes according to the growth of the underlying stock and the (gross) management fee is deducted. That is, 
        \begin{equation} \label{eq3:subaccount}
            F_t = F_{(t-1)_+}\cdot\frac{S_{t}}{S_{t-1}}\cdot(1-\eta_g)\cdot(1-q_t),
        \end{equation} 
    where $(x)^+=\max\{x,0\}$ and $F_{(t-1)_+}$ will be defined later. The insurer's income at time $t$ is the net management fee, i.e., $F_t\eta_n$.

@@

Consider a \gls{va} contract whose delta hedge portfolio at any time~$t$, $t=0,1,\ldots,T-1$, consists of $\Delta_t$ units in the underlying stock and $B_t$ amount of a risk-free zero-coupon bond maturing at time $T$.
The value of the hedge portfolio at time~$(t-1)$ is:
\begin{equation*}
    H_{t-1} = \Delta_{t-1} S_{t-1} + B_{t-1},
\end{equation*}
where $S_t$ is the underlying stock price and any time $t>0$.
This hedge portfolio is brought forward to the next rebalancing time~$t$, when its value becomes:
\begin{equation*}
    H_{t}^{bf} = \Delta_{t-1} S_{t} + B_{t-1}e^{-rt}.
\end{equation*}
Therefore, the time~$t$ hedging error, i.e., the cash flow incurred by the insurer due to rebalancing at time~$t$, is
\begin{equation}
    HE_t = H_t - H^{bf}_t, \quad t=1,\ldots, T-1.
\end{equation}
The P\&L of the \gls{va} contract includes the cost of the initial hedge ($H_0$), the hedging errors~\eqref{eq2:hedgingerror}, the unwinding of the hedge at maturity ($H^{bf}_T$), and the contract payout at time~$t\in \{0,\ldots,T\}$.
Mathematically, the present value of the liability for a \gls{gmmb} contract is 
\begin{align} \label{eq3:lossGMMB}
L   & = H_0 - e^{-rT} H^{bf}_T + \sum_{t=1}^{T-1} e^{-rt} HE_t + \sum_{t=1}^T e^{-rt} C_t - \sum_{t=1}^T e^{-rt} F_t\eta_n + e^{-rT} (G_t - F_t)^+  \nonumber \\ 
    & = e^{-rT} (G_T - F_{T^+})^+ + \sum_{t=1}^T e^{-rt}  \left( \Delta_{t-1} (S_{t-1} - e^{-r} S_t) + C_t - F_t\eta_n \right) 
\end{align}
where the equality holds by a rearrangement of terms and a telescopic sum simplification of $e^{-rt}B_t$, $t=0,\ldots,T-1$, and $C_t := C \cdot S_{t-1} \cdot |\Delta_t - \Delta_{t-1}|$ is the tranaction cost at time $t$~\citep{garleanu2013dynamic}.
Hence, the present liability consist of the hedging errors, the transaction costs, the net management fees, and the contract payouts.
Similarly, the present value of the liability for a \gls{gmwb} contract is
\begin{equation} \label{eq3:lossGMWB}
L = \sum_{t=1}^T e^{-rt}  \left( \Delta_{t-1} (S_{t-1} - e^{-r} S_t) + C_t - F_t\eta_n \right)
\end{equation}
The \gls{gmwb} payout is implicitly included in $F_t\eta_n$.
Both Equation~\eqref{eq3:lossGMMB} and Equation~\eqref{eq3:lossGMWB} neglect the transaction cost to liquidate the hedge portfolio at maturity.
The discrete-time hedging problem of \gls{va} contracts can be conveniently formulated as an \gls{mdp}, where the states, actions, transition probabilities, policies, and rewards are defined as follows.
Different from the formulation in Chapter~\ref{chap:project2}, the hedging weight $\Delta_0, \ldots, \Delta_{T-1}$ are not estimated from inner simulation steps but are directly controlled by a \gls{drl} agent.
Thus, the hedging weight $\Delta_t$ at time $t$ is estimated using a neural network whose input is the state $\mathbf{s}_t$ and output is $\Delta_t$.

@@

For hedging \gls{va} contracts,~\cite{xu2020variable} propose a deep hedging algorithm that falls into this category of model-based \gls{rl}.
A cost function is defined to minimize the expected shortfall of the total present liability.
~\cite{carbonneau2021deep} extends the model-based approach to hedging long-term financial derivatives and compares the performance of deep hedging strategies trained with different cost functions.
In deep hedging~\citep{buehler2019deep} the state space of the \gls{mdp} contains the previous hedging weight.
A more recent paper by~\cite{imaki2021no} proposes a new network architecture that removes the previous hedging weight from the state space by introducing a no-transaction band.
It allows for a more efficient training of the neural network by removing the semi-recurrent problem structure.

@@

In a model-free reinforcement learning approach, the agent is a trial-and-error learner that interacts with the environment to learn the optimal policy by iteratively updating the policy and the value function based on the observed states, actions, and rewards.
For hedging \gls{va} contracts, the agent does not have access to a planning algorithm or the underlying stochastic asset model, and it must learn the optimal policy by directly interacting with the environment.

Model-free \gls{rl} algorithms gain its popularity in robotics tasks, where transition probabilities in the environment are unknown and can be affected by the agent's actions.
In the context of hedging, existing studies use transaction costs as a function of the hedging weights, which makes the transition probabilities independent on the actions.
However, the trade size can affect the asset price in a real environment~\citep{hasbrouck1991measuring}, which makes the model-free approach a promising direction for hedging \gls{va} contracts in the real market.

@@

Representative algorithms include the \gls{trpo} algorithm proposed by~\cite{schulman2015trust} and the \gls{ppo} algorithm proposed by~\cite{schulman2017proximal}.
~\cite{du2020deep} show that the \gls{ppo} algorithm is more sample efficient and stable than the value-based approaches in training agents to hedge financial derivatives.
The most relevant paper to our study is~\cite{chong2023pseudo}, which uses the \gls{ppo} algorithm to hedge \gls{gmmb} with a \gls{gmdb} rider under the Black-Scholes framework.
However, information of the underlying risk factors is leaked to the agent, which makes their study pseudo-model-free.
To further improve the risk management of \gls{va} contracts under rough volatility models, we propose a model-free deep reinforcement learning approach to hedge \gls{va} contracts without explicit or implicit knowledge of the underlying risk factors.

\section{Hedging VA contracts with a PPO Agent} \label{sec:PPO}
In a hedging problem, an insurer aims at finding a policy $\pi$ that maximizes the expected reward over time.
In a policy-based \gls{rl} approach like the \gls{ppo}, both the policy and the value function are parameterized by deep neural networks.
We denote the policy by $\pi_{\theta}$ and the value function by $V_{\phi}$, where $\theta$ and $\phi$ are the parameters of the value network and the policy network, respectively.
The policy is updated according to an objective function that reflects the advantage of taking action $a_t$ in state $s_t$ over following the policy $\pi$.
\begin{align}
    \max_{\theta} \mathbb{E}\left[ \min \left\{ \Lambda_t(\theta)\hat{A}^{\pi_{\theta}}(s_t, a_t), \text{clip}(\Lambda_t(\theta), 1-\epsilon, 1 + \epsilon) \hat{A}^{\pi_{\theta}}(s_t, a_t)  \right\} \right] \\
    \Lambda_t(\theta) = \frac{\pi_{\theta}(a_t|s_t)}{\pi_{\theta_{old}}(a_t|s_t)} 
\end{align}
where $\theta_{old}$ denotes the policy parameters at the previous iteration, $\epsilon$ is a clip range hyperparameter that controls the size of the policy update, $\text{clip}(x, a, b)$ is a function that clips the value of $x$ to be within the range $[a, b]$, and $\hat{A}^{\pi_{\theta}}(s_t, a_t)$ is an estimated advantage of taking action $a_t$ in state $s_t$ over following the policy $\pi_{\theta_{old}}$.
In our implementation, the advantage function is estimated following the GAE method proposed by~\cite{schulman2015high}.
The policy is updated by performing gradient ascent with respect to the policy parameters $\theta$ to maximize the objective function, which is estimated by sampling from the environment and computing the empirical average of the objective function over the sampled trajectories.
The first component in the objective function, $\Lambda_t(\theta) \hat{A}^{\pi_{\theta}}(s_t, a_t)$, is a surrogate objective function that approximates the advantage of taking action $a_t$ in state $s_t$ over following the policy $\pi_{\theta_{old}}$, and the second component, $\text{clip}(\Lambda_t(\theta), 1-\epsilon, 1 + \epsilon) \hat{A}^{\pi_{\theta}}(s_t, a_t)$, is a clipped surrogate objective function that prevents large policy updates and ensures more stable training.
The \gls{trpo} algorithm proposed by~\cite{schulman2015trust} uses a similar clipped surrogate objective function to ensure that the policy update does not deviate too far from the old policy, but \gls{ppo} simplifies the optimization problem by using a clipped surrogate objective function instead of a hard-coded constraint on the policy update.
Compared to \gls{trpo}, \gls{ppo} is more computationally efficient and easier to implement, making it a popular choice for training deep reinforcement learning agents.
For a in-depth discussion on the implementation details of both \gls{trpo} and \gls{ppo}, we refer the reader to~\cite{engstrom2020implementation}.

@@

Algorithm~\ref{alg3:ppoHedging} outlines the \gls{ppo} algorithm for training an agent to hedge \gls{va} contracts.
At each iteration $k$, the agent collects a set of trajectories $\mathcal{D}_k$ by interacting with the environment (Section~\ref{subsec:VASimulation}) with the current policy $\pi_{\theta_k}$.
The agent then computes the rewards $\hat{R}_t$ for each trajectory in $\mathcal{D}_k$ and estimates the advantage function with $\hat{A}_{\pi_{\theta_k}}(s_t, a_t)$ using the value function $V_{\phi_k}$.
The policy parameters $\theta_{k+1}$ are updated by performing gradient ascent with respect to the policy parameters to maximize the clipped surrogate objective function.
The value function parameters $\phi_{k+1}$ are updated by minimizing the mean squared error between the predicted value and the actual collected reward.
The agent iterates this process until convergence to learn an optimal policy for hedging \gls{va} contracts.

@@

In the context of hedging \gls{va} contracts, the state space is partially observed.
The agent does not have access to the contract specifications and the underlying risk factors.
Furthermore, a rough volatility model introduces non-Markovian dynamics, i.e., the future state of the system depends on the entire history of the system.
To address the non-Markovian dynamics, an \gls{lstm} component is added to the state representation $\mathbf{s}_t$ to capture the temporal dependencies in the data.
This approach is known as recurrent \gls{ppo}~\citep{ni2021recurrent}.
Let $h_t$ be the hidden state of the \gls{lstm} at time $t$.
The state representation $\mathbf{s}_t$ is fed into the \gls{lstm} at each time step to update the hidden state $h_t$.
\begin{equation} \label{eq3:stateRNN}
    \mathbf{s}_t = (S_t, F_t, G_t, \tau, \Delta_{t-1}), \quad h_t = \text{LSTM}(\mathbf{s}_t, h_{t-1}),
\end{equation}
where 
\begin{align} 
    h_t &= o_t \odot \tanh(c_t), \nonumber \\
    c_t &= f_t \odot c_{t-1} + i_t \odot g_t, \nonumber \\
    i_t &= a(W_i \mathbf{s}_t + U_i h_{t-1} + b_i), \nonumber \\
    f_t &= a(W_f \mathbf{s}_t + U_f h_{t-1} + b_f), \nonumber \\
    o_t &= a(W_o \mathbf{s}_t + U_o h_{t-1} + b_o), \nonumber \\
    g_t &= a(W_g \mathbf{s}_t + U_g h_{t-1} + b_g), 
\end{align}
In a \gls{lstm}, the cell state $c_t$ stores a representation of historical information up to time $t$.
It is updated at each time step by a forget gate $f_t$, an input gate $i_t$, and an output gate $o_t$.
Equation~\ref{eq:lstm} provides a general intoduction to \gls{lstm} for a generic time series input $x_t$.

In a recurrent \gls{ppo} with \gls{lstm}, the policy and reward function remain the same but with the new state representation in Equation~\eqref{eq3:stateRNN}.
Instead of following the standard \gls{ppo} algorithm, the agent uses the recurrent \gls{ppo} algorithm to learn an optimal policy for hedging \gls{va} contracts.
Figure~\ref{fig3:architecturePPO} illustrates the network architecture of the standard \gls{ppo} and the recurrent \gls{ppo} with \gls{lstm}, and Algorithm~\ref{alg3:ppoHedging-rnn} outlines the recurrent \gls{ppo} algorithm and for training an agent to hedge \gls{va} contracts with non-Markovian dynamics.

@@

\section{Transfer Learning for Hedging VA contracts} \label{sec:TransferLearning}

In the context of deep hedging, various research has been conducted to improve the hedging strategies on real-world data by leveraging knowledge from simulated data.
Most of the research uses the offline-online learning approach, where the hedging strategies are first trained with simulation data, but updated in real-time based on new market information and continuous learning to adapt to changing market dynamics.
The most relevant paper to our study is~\cite{xiao2021optimal}, which use policy gradient algorithms to train hedging strategies for European options with simulation data from Heston model and then updated in real-time hedging S\&P 500 index options.
To the best of our knowledge, no research has been conducted to improve the hedging strategies with the state-of-the-art transfer learning techniques.
This section introduces some transfer learning techniques that are relevant to our task of hedging \gls{va} contracts.

@@

In the context of hedging \gls{va} contracts using deep reinforcement learning and transfer learning, reward shaping becomes a crucial technique. 
It involves modifying the reward function in the target domain to incorporate additional guidance, which often leads to more efficient learning and a better performance.
This technique can be particularly useful for managing the complex risks associated with \gls{va} contracts, as it can guide the learning process towards strategies that are more effective in hedging these products.

@@

We present a numerical experiment to demonstrate the effectiveness of the model-free \gls{ppo} algorithm for hedging \gls{va} contracts with transaction costs.
We consider a \gls{va} contract with a maturity of $T=240$ months and a \gls{gmmb} rider with a simulation environment described in Section~\ref{subsec:VASimulation}.
The simulation enviornment is implemented in Python using the OpenAI Gym framework~\citep{brockman2016openai}, and the neural networks are programmed using the PyTorch library~\citep{paszke2019pytorch}.
The environment is designed to simulate the dynamics of the \gls{va} contract and provide the agent with the current state of the \gls{va} with the contract specifications and the financial market information described in Table~\ref{tab3:hyperparameters}.

@@

Figure~\ref{fig3:dh-transfers} illustrates the hedging performance of the deep hedging algorithm with transfer learning for hedging \gls{va} contracts with a \gls{gmmb} rider.
The first deep hedging agent is directly trained on the Heston asset model, and the second deep hedging agent is trained with transfer learning from the \gls{gbm} model.
The figure plot the 95\%-CVaR of hedging errors against training epochs for both agents.
The deep hedging agent with transfer learning outperforms the deep hedging agent trained on the Heston model in both mastery and generalization.
The agent with transfer learning achieves a lower loss in the target domain and adapts more quickly to the new environment, indicating that the agent has learned a more robust hedging strategy that can be effectively applied to a more complex environment.

@@

This experiment compares the hedging performance of the recurrent \gls{ppo} agent with \gls{lstm} and the deep hedging algorithm proposed by~\cite{buehler2019deep} for hedging \gls{va} contracts with a \gls{gmmb} rider.
The experiment is conducted with the same amount ofsimulation budget.
